---
title: "Geometric Brownian Motion Simulation"
editor_options: 
  chunk_output_type: console
---

# `R` Euler-Maruyama discretization:

Since there is a closed-form solution (below) for the GBM stochastic differential equation, this scheme is not necessary, but a nice illustration nevertheless.

For an Ito process, $X=(X)_{t>t_0}$, that satisfies the SDE:

$$
X_t = x + \int_{t_0}^t\mu(s,X_s)ds + \int_{t_0}^t\sigma(s,X_s)dW_s
$$

$$
dX_t = \mu(t,X_t)dt + \sigma(t,X_t)dW_t
$$

The Euler-Maruyama discretisation can be implemented recursively as:

$$
X_{t_{n+1}} = X_{t_n} + \mu(t,X_{t_n})\Delta t + \sigma(t_n,X_{t_n})\epsilon_n
\qquad 
\epsilon \sim N(0,\Delta t)
$$

```{r}
GBM_Euler <- function(S0, r, v, t, N = 365) {
    dt      <- t / N
    time    <- seq(from = 0, to = t, by = dt)
    
    Z       <- rnorm(n = N, mean = 0, sd = 1)
    dW      <- sqrt(dt) * Z
    W       <- c(0, cumsum(dW))
    St      <- c(S0, numeric(N))

    for (i in 2:length(St)) {
        St[i] <- St[i-1] + r * St[i-1] * dt + v * St[i-1] * dW[i-1]
    }
    
    return( St )
}
```

# `C++` Closed-Form Simulation:

```{Rcpp, file = "GBM.cpp"}

```

```{r}
Rcpp::sourceCpp(file = "GBM.cpp")
```

# `R` Closed-Form Simulation:

```{r}
GBM_R <- function(
        # FUNCTION ARGUMENTS
        S0,       # initial stock value
        t,        # time to expiry
        r,        # risk-free rate
        v,        # volatility 
        N = 1000  # number of steps
        ) {
    
    dt = t / N                # length of each step
    time = seq(0, t, dt)      # vector of time points
    
    Z   <- rnorm(N)             # standard normal sample of N elements
    dW  <- Z * sqrt(dt)         # Brownian motion increments
    W   <- c(0, cumsum(dW))     # Brownian motion at each time (N+1 elements)
    
    St  <- c(S0, numeric(N))    # vector of zeros, N+1 elements
    St  <-  S0 * exp((r - v^2 / 2) * time + v * W)
    return( St )
}
```

# Comparison:

```{r}
R_sims_euler <- replicate(
    n = 20,
    GBM_Euler(t = 1, r = 0.05, v = 0.3, S0 = 100, N = 1000)
)

R_sims_analytic <- replicate(
    n = 20,
    GBM_R(t = 1, r = 0.05, v = 0.3, S0 = 100, N = 1000)
)

Cpp_sims <- replicate(
    n = 20,
    GBM_Cpp(tau = 1, r = 0.05, sigma = 0.3, S0 = 100, N = 1000)
)

matplot(R_sims_euler, type = "l", main="R Simulations (Euler-Maruyama)")
matplot(R_sims_analytic, type = "l", main="R Simulations (analytic)")
matplot(Cpp_sims, type = "l", main="C++ Simulations")

microbenchmark::microbenchmark(
    GBM_Euler = GBM_Euler(S0 = 100, r = 0.05, v = 0.2, t = 1, N = 365),
    GBM_R = GBM_R(t = 1, r = 0.05, v = 0.2, S0 = 100, N = 365),
    GBM_Cpp = GBM_Cpp(tau = 1, r = 0.05, sigma = 0.2, S0 = 100, N = 365),
    times = 10000
)
```






