---
title: "Asian Options"
editor_options: 
  chunk_output_type: console
---

Asian options are options that depend on the average price of the underlying over the life of the option. One can use either the geometric mean ($\bar{x}=\sqrt[n]{x_1 \cdots x_n}$), or the arithmetic mean ($\bar{x} = \frac{1}{n}(x_1+\cdots + x_n)$).

Closed form solutions exist for geometric mean Asian options since the mean itself is also log-normally distributed, however for arithmetic mean Asian options, generally approximation formulas or Monte Carlo simulations are required. 

# Types of Asian options

## Average Price Option

+ Average Price Call Payoff
    
$$
(A(0,T) - K)^+ \qquad \text{or} \qquad (G(0,T) - K)^+
$$

+ Average Price Put Payoff

$$
(K - A(0,T))^+ \qquad \text{or} \qquad (K - G(0,T))^+
$$


## Average Strike Option

+ Average Strike Call Payoff
    
$$
(S(0,T) - A(0,T))^+ \qquad \text{or} \qquad (S(0,T) - G(0,T))^+
$$

+ Average Strike Put Payoff
    
$$
(A(0,T) - S(0,T))^+ \qquad \text{or} \qquad (G(0,T) - S(0,T))^+
$$


Where the types of averaging are as follows (continuous vs. discrete):

$$
A(0,T) = \frac{1}{T}\int_0^T S(t)dt 
\qquad \qquad
A(0,T) = \frac{1}{n}\sum_{i=1}^n S(t_i)
$$

$$
G(0,T) = \exp\left(\frac{1}{T}\int_0^T \ln[S(t)] dt \right)
\qquad \qquad
G(0,T) = \sqrt[n]{\prod{S(t_i)}} = \exp\left(\frac{1}{n}\sum_{i=1}^n\ln[S(t_i)]\right)
$$



# Geometric Average

## Closed-form solution

A geometric continuous average-rate option can be priced by making slight adjustments to the pricing formulas for standard options as follows (Kemna & Vorst, 1990):

$$
C = Se^{(b_A - r)T}N(d_1) - Xe^{-rT}N(d_2)
$$

$$
P = - Se^{(b_A - r)T}N(-d_1) + Xe^{-rT}N(-d_2)
$$

Adjustments to $d_1$, $d_2$, volatility and cost-of-carry:

$$
\begin{aligned}
d_1 &= \frac{\ln(S/X) + (b_A + \sigma_A^2/2)T}{\sigma_A\sqrt{T}} \\
d_2 &= d_1 - \sigma_A\sqrt{T} \\
\sigma_A &= \frac{\sigma}{\sqrt{3}} \\
b_A &= \frac{1}{2}\left( b - \frac{\sigma^2}{6} \right)
\end{aligned}
$$


```{r}
AsianGeometricClosedForm <- function(S, X, r, t, v, q = 0.0) {
    
    b <- r - q
    
    vA <- v / sqrt(3)
    bA <- 0.5 * (b - v^2/6)
    
    d1 <- (log(S/X) + (bA + vA^2/2) * t) / (vA * sqrt(t))
    d2 <- d1 - vA * sqrt(t)
    
    prices <- list(
        "Call price" = S*exp((bA-r)*t)*pnorm(d1) - X*exp(-r*t)*pnorm(d2),
        "Put price" = -S*exp((bA-r)*t)*pnorm(-d1) + X*exp(-r*t)*pnorm(-d2)
    )
    
    return( prices )
}
```

## Monte Carlo simulation

For the Monte Carlo method we can use the Geometric Brownian Motion simulation `C++` file [here](GBM.html).

```{r}
Rcpp::sourceCpp(file = "GBM.cpp")
```

```{r}
AsianGeometricMonteCarlo <- function(
        S, K, r, t, v, 
        q = 0.0,
        paths = 1000L,
        time_steps = 1000L) {
    
    # browser()
    
    # print("Simulating Geometric Brownian Motion:")
    # S <- plyr::raply(
    #     .n = paths,
    #     .expr = GBM_Cpp(tau = t, N = time_steps, r = r, sigma = v, S0 = S),
    #     .progress = "text"
    # )
    # 
    # print("Calculating geometric mean of GBM paths:")
    # means <- plyr::aaply(
    #     .data = S,
    #     .margins = 1,
    #     .fun = function(S) exp(mean(log(S))),
    #     .progress = "text"
    # )
    
    S <- replicate(
        n = paths,
        expr = GBM_Cpp(tau = t, N = time_steps, r = r, sigma = v, S0 = S)
    )

    means <- apply( S, 2, function(S) exp(mean(log(S))) )
    
    prices <- list(
        "Call price" = mean( exp( -r*t) * pmax(means - K, 0) ) ,
        "Put price" = mean( exp( -r*t) * pmax(K - means, 0) ) 
    )
    
    return( prices )
}
```

## Comparison of geometric mean methods

```{r}
AsianGeometricClosedForm(S = 100, X = 110, r = 0.05, t = 1, v = 0.2)
AsianGeometricMonteCarlo(S = 100, K = 110, r = 0.05, t = 1, v = 0.2, time_steps = 10000L, paths = 1000L)
```


# Arithmetic Average 

## Turnbull-Wakeman (1991) approximation





