---
title: "Barrier Options"
editor_options: 
  chunk_output_type: console
---


# Closed-form solutions:

Barrier options are path-dependent, as their payoffs depend on whether or not the underlying price hits a particular boundary up until, and including, the maturity.

There are closed-form pricing formulas for these options [Merton (1973); Reiner & Rubenstein (1991); Haug (2006)]. For the closed-form solutions we need (from *Option Pricing Formulas* by Espen Haug, 2006):


$$
\begin{aligned}
& A = \phi S e^{(b-r)T}N(\phi x_1) - \phi Xe^{-rT}N(\phi x_1 - \phi \sigma \sqrt{T}) \\
& B = \phi S e^{(b-r)T}N(\phi x_2) - \phi Xe^{-rT}N(\phi x_2 - \phi \sigma \sqrt{T}) \\
& C = \phi Se^{(b-r)T}(H/S)^{2(\mu+1)}N(\eta y_1) - \phi Xe^{-rT}(H/S)^{2\mu}N(\eta y_1 - \eta \sigma \sqrt{T}) \\
& D = \phi Se^{(b-r)T}(H/S)^{2(\mu+1)}N(\eta y_2) - \phi Xe^{-rT}(H/S)^{2\mu}N(\eta y_2 - \eta \sigma \sqrt{T}) \\
& E = Ke^{-rT}[N(\eta x_2 - \eta \sigma \sqrt{T}) - (H/S)^{2\mu}N(\eta y_2 - \eta \sigma \sqrt{T})] \\
& F = K[(H/S)^{\mu+\lambda}N(\eta z) + (H/S)^{\mu-\lambda}N(\eta z - 2\eta\lambda\sigma\sqrt{T})] \\
\end{aligned}
$$


Where:


$$
x_1 = \frac{\ln(S/X)}{\sigma\sqrt{T}} + (1+\mu)\sigma\sqrt{T}
\qquad
x_2 =  \frac{\ln(S/H)}{\sigma\sqrt{T}} + (1+\mu)\sigma\sqrt{T}
$$


$$
y_1 = \frac{\ln(H^2/(SX))}{\sigma\sqrt{T}} + (1+\mu)\sigma\sqrt{T}
\qquad
y_2 =  \frac{\ln(H/S)}{\sigma\sqrt{T}} + (1+\mu)\sigma\sqrt{T}
$$


$$
z = \frac{\ln(H/S)}{\sigma\sqrt{T}} + \lambda\sigma\sqrt{T}
\qquad
\mu =  \frac{b-\sigma^2/2}{\sigma^2}
\qquad
\lambda = \sqrt{\mu^2 + \frac{2r}{\sigma^2}}
$$


## Knock-in:

### Down-and-In call: 

$$
\text{Payoff} = 
\begin{cases}
(S-X)^+ & \text{if }S \leq H \text{ before T} \\
K & \text{else}
\end{cases}
$$

$$
\begin{aligned}
C_{DI} = (C + E) \cdot \mathbf{1}_{X>H} + (A-B+D+E)\cdot \mathbf{1}_{X<H}
&& \eta=1, \phi=1
\end{aligned}
$$

### Up-and-In call: 

$$
\text{Payoff} = 
\begin{cases}
(S-X)^+ & \text{if }S \geq H \text{ before T} \\
K & \text{else}
\end{cases}
$$


$$
\begin{aligned}
C_{UI} = (A + E) \cdot \mathbf{1}_{X>H} + (B-C+D+E)\cdot \mathbf{1}_{X<H}
&& \eta=-1, \phi=1
\end{aligned}
$$

### Down-and-In put: 

$$
\text{Payoff} = 
\begin{cases}
(X-S)^+ & \text{if }S \leq H \text{ before T} \\
K & \text{else}
\end{cases}
$$


$$
\begin{aligned}
P_{DI} = (B-C+D+E) \cdot \mathbf{1}_{X>H} + (A+E)\cdot \mathbf{1}_{X<H}
&& \eta=1, \phi=-1
\end{aligned}
$$

### Up-and-In put: 

$$
\text{Payoff} = 
\begin{cases}
(X-S)^+ & \text{if }S \geq H \text{ before T} \\
K & \text{else}
\end{cases}
$$


$$
\begin{aligned}
P_{UI} = (A-B+D+E) \cdot \mathbf{1}_{X>H} + (C+E)\cdot \mathbf{1}_{X<H}
&& \eta=-1, \phi=-1
\end{aligned}
$$

## Knock-Out:

### Down-and-Out call: 

$$
\text{Payoff} = 
\begin{cases}
(S-X)^+ & \text{if } S > H \text{ before T} \\
K & \text{else (at hit)}
\end{cases}
$$

$$
\begin{aligned}
C_{DO} = (A-C+F) \cdot \mathbf{1}_{X>H} + (B-D+F)\cdot \mathbf{1}_{X<H}
&& \eta=1, \phi=1
\end{aligned}
$$

### Up-and-Out call: 

$$
\text{Payoff} = 
\begin{cases}
(S-X)^+ & \text{if } S < H \text{ before T} \\
K & \text{else (at hit)}
\end{cases}
$$


$$
\begin{aligned}
C_{UO} = F \cdot \mathbf{1}_{X>H} + (A-B+C-D+F)\cdot \mathbf{1}_{X<H}
&& \eta=-1, \phi=1
\end{aligned}
$$

### Down-and-Out put: 

$$
\text{Payoff} = 
\begin{cases}
(X-S)^+ & \text{if }S > H \text{ before T} \\
K & \text{else (at hit)}
\end{cases}
$$


$$
\begin{aligned}
P_{DO} = (A-B+C-D+F) \cdot \mathbf{1}_{X>H} + F\cdot \mathbf{1}_{X<H}
&& \eta=1, \phi=-1
\end{aligned}
$$

### Up-and-Out put: 

$$
\text{Payoff} = 
\begin{cases}
(X-S)^+ & \text{if }S < H \text{ before T} \\
K & \text{else (at hit)}
\end{cases}
$$


$$
\begin{aligned}
P_{UO} = (B-D+F) \cdot \mathbf{1}_{X>H} + (A-C+F)\cdot \mathbf{1}_{X<H}
&& \eta=-1, \phi=-1
\end{aligned}
$$



```{r}
BarrierOptionClosedForm <- function(
        # Function arguments:
        S, # Underlying price
        X, # Strike price
        H, # Barrier 
        t, # Time to expiry
        r, # Risk-free rate
        v, # Volatility
        K = 0.0, # Rebate (optional)
        q = 0.0  # Dividend yield (optional)
        ) {

    # Cost of carry = risk-free rate - dividend yield
    b       <- r - q

    # Common terms: 
    mu      <- (b - v^2 / 2) / v^2
    lambda  <- sqrt(mu^2 + 2*r / v^2)
    z       <- log(H/S) / (v*sqrt(t)) + lambda * v * sqrt(t)
    
    x1      <- log(S / X) / v*sqrt(t) + (1 + mu) * v * sqrt(t)
    x2      <- log(S / H) / v*sqrt(t) + (1 + mu) * v * sqrt(t)
    
    y1      <- log(H^2 / (S * X)) / (v*sqrt(t)) + (1 + mu) * v * sqrt(t)
    y2      <- log(H / S) / (v*sqrt(t)) + (1 + mu) * v * sqrt(t)
    
    # Common functions: 
    AA <- function(phi) {
        term1 <- phi * S * exp( (b-r) * t ) * pnorm(phi * x1)
        term2 <- phi * X * exp( -r * t ) * pnorm(phi * x1 - phi * v * sqrt(t))
        return( term1 - term2 )
    }
    
    BB <- function(phi) {
        term1 <- phi * S * exp( (b - r) * t ) * pnorm(phi * x2)
        term2 <- phi * X * exp( -r * t ) * pnorm(phi * x2 - phi * v * sqrt(t))
        return( term1 - term2 )
    }
    
    CC <- function(phi, eta) {
        term1 <- phi * S * exp( (b - r) * t ) * (H/S)^(2 * (mu + 1)) * pnorm(eta * y1)
        term2 <- phi * X * exp( -r * t ) * (H/S)^(2 * mu) * pnorm(eta * y1 - eta * v * sqrt(t))
        return( term1 - term2 )
    }
    
    DD <- function(phi, eta) {
        term1 <- phi * S * exp( (b - r) * t ) * (H/S)^(2 * (mu + 1)) * pnorm(eta * y2)
        term2 <- phi * X * exp( -r * t ) * (H/S)^(2 * mu) * pnorm(eta * y2 - eta * v * sqrt(t))
        return( term1 - term2 )
    }
    
    EE <- function(eta) {
        term1 <- pnorm(eta * x2 - eta * v * sqrt(t))
        term2 <- (H/S)^(2 * mu) * pnorm(eta * y2 - eta * v * sqrt(t))
        return( K * exp( -r * t ) * (term1 - term2) )
    }
    
    FF <- function(phi, eta) {
        term1 <- (H/S)^(mu + lambda) * pnorm(eta * z)
        term2 <- (H/S)^(mu - lambda) * pnorm(eta * z - 2 * eta * lambda * v * sqrt(t))
        return( K * (term1 + term2) )
    }
    
    # Prices dataframe:
    PRICES <- data.frame(
        "Down-and-Out" = c(NA, NA), 
        "Down-and-In" = c(NA, NA), 
        "Up-and-Out" = c(NA, NA), 
        "Up-and-In" = c(NA, NA),
        row.names = c("Call Price", "Put Price")
    )
    
    # Strike above barrier:
    if ( X >= H ) {
        # Knock-In calls:
        PRICES$Down.and.In[1]  <- ifelse(S<=H, "S<=H", CC(1, 1) + EE(1)) 
        PRICES$Up.and.In[1]    <- ifelse(S>=H, "S>=H", AA(1) + EE(-1)) 
        # Knock-In puts:
        PRICES$Down.and.In[2]  <- ifelse(S<=H, "S<=H", BB(-1) - CC(-1, 1) + DD(-1, 1) + EE(1)) 
        PRICES$Up.and.In[2]    <- ifelse(S>=H, "S>=H", AA(-1) - BB(-1) + DD(-1, -1) + EE(-1)) 
        # Knock-Out calls:
        PRICES$Down.and.Out[1] <- ifelse(S<=H, "S<=H", AA(1) - CC(1, 1) + FF(1, 1)) 
        PRICES$Up.and.Out[1]   <- ifelse(S>=H, "S>=H", FF(1, -1)) 
        # Knock-Out puts:
        PRICES$Down.and.Out[2] <- ifelse(S<=H, "S<=H", AA(-1) - BB(-1) + CC(-1, 1) - DD(-1, 1) + FF(-1, 1)) 
        PRICES$Up.and.Out[2]   <- ifelse(S>=H, "S>=H", BB(-1) - DD(-1, -1) + FF(-1, -1)) 
    }
    
    # Strike below barrier:
    if ( X < H ) {
        # Knock-In calls:
        PRICES$Down.and.In[1]  <- ifelse(S<=H, "S<=H", AA(1) - BB(1) + DD(1, 1) + EE(1)) 
        PRICES$Up.and.In[1]    <- ifelse(S>=H, "S>=H", BB(1) - CC(1, -1) + DD(1, -1) + EE(-1)) 
        # Knock-In puts:
        PRICES$Down.and.In[2]  <- ifelse(S<=H, "S<=H", AA(-1) + EE(1)) 
        PRICES$Up.and.In[2]    <- ifelse(S>=H, "S>=H", CC(-1, -1) + EE(-1)) 
        # Knock-Out calls:
        PRICES$Down.and.Out[1] <- ifelse(S<=H, "S<=H", BB(1) - DD(1, 1) + FF(1, 1)) 
        PRICES$Up.and.Out[1]   <- ifelse(S>=H, "S>=H", AA(1) - BB(1) + CC(1, -1) - DD(1, -1) + FF(1, -1)) 
        # Knock-Out puts:
        PRICES$Down.and.Out[2] <- ifelse(S<=H, "S<=H", FF(-1, 1)) 
        PRICES$Up.and.Out[2]   <- ifelse(S>=H, "S>=H", AA(-1) - CC(-1, -1) + FF(-1, -1)) 
    }
    
    return( PRICES )
}
```



# Monte-Carlo Pricing:

An alternative is to use Monte-Carlo simulation to generate Geometric Brownian Motion (GBM) trajectories of a stock price and estimate the price of the barrier option based on these samples. To do so, we create a function to simulate the GBM trajectories:


```{Rcpp, file = "GBM.cpp"}

```


```{r}
Rcpp::sourceCpp(file = "GBM.cpp")
```


```{r}
BarrierOptionMonteCarlo <- function(
        S, X, r, t, v, H, 
        K = 0.0,
        q = 0.0, 
        paths = 1000L, 
        time_steps = 1000L) {
    
    # browser()
    
    S0 <- S
    
    # Generate Geometric Brownian Motion samples
    S <- replicate(
        n = paths,
        expr = GBM_Cpp(tau = t, N = time_steps, r = r, sigma = v, S0 = S)
    )
    
    P <- data.frame(
        "DO" = ifelse( c(S0, S0) > H , c(NA, NA) , c("S<=H", "S<=H") ),
        "DI" = ifelse( c(S0, S0) > H , c(NA, NA) , c("S<=H", "S<=H") ),
        "UO" = ifelse( c(S0, S0) < H , c(NA, NA) , c("S>=H", "S>=H") ),
        "UI" = ifelse( c(S0, S0) < H , c(NA, NA) , c("S>=H", "S>=H") ),
        row.names = c("Call Price", "Put Price")
    )
    
    # Knock-In calls:
    if ( is.na(P$DI[1]) ) {
        P$DI[1] <- sum(apply(S, 2, function(S) ifelse(all(S > H), K, max(S[length(S)] - X, 0)))) / paths
    }
    if ( is.na(P$UI[1]) ) {
        P$UI[1] <- sum(apply(S, 2, function(S) ifelse(all(S < H), K, max(S[length(S)] - X, 0)))) / paths
    }
    # Knock-In puts:
    if ( is.na(P$DI[2]) ) {
        P$DI[2] <- sum(apply(S, 2, function(S) ifelse(all(S > H), K, max(X - S[length(S)], 0)))) / paths
    }
    if ( is.na(P$UI[2]) ) {
        P$UI[2] <- sum(apply(S, 2, function(S) ifelse(all(S < H), K, max(X - S[length(S)], 0)))) / paths
    }
    # Knock-Out calls:
    if ( is.na(P$DO[1]) ) {
        P$DO[1] <- sum(apply(S, 2, function(S) ifelse(any(S <= H), K, max(S[length(S)] - X, 0)))) / paths
    }
    if ( is.na(P$UO[1]) ) {
        P$UO[1] <- sum(apply(S, 2, function(S) ifelse(any(S >= H), K, max(S[length(S)] - X, 0)))) / paths
    }
    # Knock-Out puts:
    if ( is.na(P$DO[2]) ) {
        P$DO[2] <- sum(apply(S, 2, function(S) ifelse(any(S <= H), K, max(X - S[length(S)], 0)))) / paths
    }
    if ( is.na(P$UO[2]) ) {
        P$UO[2] <- sum(apply(S, 2, function(S) ifelse(any(S >= H), K, max(X - S[length(S)], 0)))) / paths
    }
    
    colnames( P ) <- c("Down.and.Out", "Down.and.In", "Up.and.Out", "Up.and.In")
    
    return( P )
}
```

# Comparing the output:

```{r}
BarrierOptionClosedForm(
    S = 100.0, # Underlying price
    X = 110.0, # Strike price
    H = 80.00, # Barrier 
    t = 01.00, # Time to expiry
    r = 0.050, # Risk-free rate
    v = 0.200 # Volatility
)

BarrierOptionMonteCarlo(
    S = 100.0, # Underlying price
    X = 110.0, # Strike price
    H = 80.00, # Barrier 
    t = 01.00, # Time to expiry
    r = 0.050, # Risk-free rate
    v = 0.200, # Volatility
    paths = 1000L,
    time_steps = 10000L
)

BarrierOptionClosedForm(
    S = 100.0, # Underlying price
    X = 110.0, # Strike price
    H = 120.0, # Barrier 
    t = 01.00, # Time to expiry
    r = 0.050, # Risk-free rate
    v = 0.200 # Volatility
)

BarrierOptionMonteCarlo(
    S = 100.0, # Underlying price
    X = 110.0, # Strike price
    H = 120.0, # Barrier 
    t = 01.00, # Time to expiry
    r = 0.050, # Risk-free rate
    v = 0.200, # Volatility
    paths = 1000L,
    time_steps = 10000L
)
```
