---
title: "Maximum Likelihood Estimation"
---

# Generalised Extreme Value Distribution for River Flow Data

We have a small dataset obtained from a 1984 consulting session on the maximum flow of rivers measured from one station ($n=35$ yearly maximums).

We want to find the MLEs of the paramaters of the location-scale extreme value distribution model, and use a QQ-Plot to check the goodness-of-fit. 

## The Maximum Likelihood Estimators

We are given the Generalised Extreme-Value Distribution, for the special case where shape parameter $\xi = 0$, which implies exponential tail decay. The density is:

$$
f(x|\mu,\sigma) 
= \frac{1}{\sigma}f_0\left(\frac{x-\mu}{\sigma}\right)
= \frac{1}{\sigma}\exp\left[\frac{\mu-x}{\sigma} - \exp\left(\frac{\mu-x}{\sigma}\right)\right]
$$

$$
f_0(x) = e^{-x}e^{-e^{-x}}, \qquad \mu \in \mathbb{R}, \qquad \sigma > 0
$$

The associated likelihood and log-likelihood functions are:

$$
\mathcal{L}(\mu,\sigma|x) 
= \prod_{i=1}^n \frac{1}{\sigma}f_0\left(\frac{x_i-\mu}{\sigma}\right)
= \prod_{i=1}^n \frac{1}{\sigma}\exp\left[\frac{\mu-x_i}{\sigma} - \exp\left(\frac{\mu-x_i}{\sigma}\right)\right]
= \frac{1}{\sigma^n}\prod_{i=1}^n \exp\left[\frac{\mu-x_i}{\sigma} - \exp\left(\frac{\mu-x_i}{\sigma}\right)\right]
$$

$$
\ln[\mathcal{L}(\mu,\sigma|x)]
= -n\ln(\sigma) + \sum_{i=1}^n \left[\frac{\mu-x_i}{\sigma} - \exp\left(\frac{\mu-x_i}{\sigma}\right)\right]
= -n\ln(\sigma) + \sum_{i=1}^n \left[\frac{\mu-x_i}{\sigma}\right] - \sum_{i=1}^n \exp\left(\frac{\mu-x_i}{\sigma}\right)
$$

This function has no closed-form solution that can be derived analytically, so we need to use use `R` to estimate the parameters numerically. 

Note that we return $-\ln[\mathcal{L}(\mu,\sigma|x)]$ since `R`'s `optim` function default method is minimising, and we are after the maximum.

```{r, cache = T}
# SET WORKING DIRECTORY TO LOCATION THIS FILE AND RIVERS.TXT ARE SAVED IN:
# setwd(dirname(rstudioapi::getSourceEditorContext()$path))
setwd(getwd())

# DECLARE AND DEFINE GEVD DENSITY AND LOG-LIKELIHOOD FUNCTIONS:
GEVD_FN <- function(parameters, x) {
    mu      <- parameters[1]
    sigma   <- parameters[2]
    FN      <- (1/sigma) * exp((mu-x)/sigma - exp((mu-x)/sigma))
    return( FN )
}

GEVD_LL <- function(parameters, x) {
    n       <- length(x)
    mu      <- parameters[1]
    sigma   <- parameters[2]
    LL      <- -n*log(sigma) + sum((mu-x)/sigma) - sum(exp((mu-x)/sigma))
    return( - LL )
}

# READ IN RIVER FLOW DATA FROM TEXT FILE:
RIVERS <- read.delim("rivers.txt", header = FALSE)[,1]
mean(RIVERS)
sd(RIVERS)

# OPTIMISE LOG-LIKELIHOOD FUNCTION FOR THE PARAMETERS:
optim(par = c(mean(RIVERS), sd(RIVERS)), 
      fn = GEVD_LL, 
      x = RIVERS, 
      method = "BFGS")

# GET PARAMETERS:
PARS <- optim(par = c(mean(RIVERS), sd(RIVERS)), 
              fn = GEVD_LL, 
              x = RIVERS, 
              method = "BFGS")$par

# PLOT THE RIVER FLOW DENSITY:
plot(density(RIVERS), 
     main = "Yearly Maximum River Flow Density", 
     xlab = "Yearly Max Flow of Rivers",
     col = "blue")
```


## QQ-Plot to Graphically Assess Goodness-of-Fit

We need the GEVD distribution function, obtained by integrating the density:

$$
F_X(x|\mu,\sigma) 
= \int_{-\infty}^x \frac{1}{\sigma}\exp\left[\frac{\mu-x}{\sigma} - \exp\left(\frac{\mu-x}{\sigma}\right)\right] dx
= \exp\left[ - \exp\left( \frac{\mu-x}{\sigma} \right) \right] 
$$

By inverting the CDF we obtain the quantile function, which we can use for our QQ-plot:

$$
Q(p|\mu,\sigma)
= F_X^{-1}(x|\mu,\sigma) 
= \mu - \sigma \ln(-\ln(p))
\qquad p \in (0,1)
$$

We can now implement the distribution and quantile functions in `R`, and obtain the QQ-plot:

```{r, cache = T}
GEVD_DF <- function(parameters, x) {
    mu      <- parameters[1]
    sigma   <- parameters[2]
    return( exp( - exp((mu-x)/sigma) ) )
}

GEVD_QF <- function(parameters, p) {
    mu      <- parameters[1]
    sigma   <- parameters[2]
    return( mu - sigma * log( - log(p) ) )
}

par(mfrow=c(1,2))
curve(GEVD_DF(PARS, x), xlim = c(0, 12000), 
      main = "Distribution Function", 
      col = "blue", xlab = "x", ylab = "")
curve(GEVD_QF(PARS, x), main = "Quantile Function",
      xlab = "p", ylab = "",
      col = "blue")

par(mfrow = c(1,1))
qqplot(GEVD_QF(PARS, seq(0, 1, 1/36))[2:36], RIVERS,
       main = "QQ-Plot to Check Goodness-of-Fit",
       xlab = "Estimated Quantiles",
       ylab = "River Flow Data Quantiles", 
       col = "blue")
qqline(RIVERS, distribution = function(p) GEVD_QF(PARS, p), col = "red")
```

